Deep ordinal classification based on cumulative link models

### 2019-05-21 paper/ieee.tex
New gauss version
Plot gauss, clmgauss parameters adjustments, clm cauchit added
Correciones paper v2

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

@@

\begin{document}
	%
	% paper title
	% Titles are generally capitalized except for words such as a, an, and, as,
	% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
	% not capitalized unless they are the first or last word of the title.
	% Linebreaks \\ can be used within to get better formatting as desired.
	% Do not put math or special symbols in the title.
	\title{Classification of ordinal data in deep learning: an experimental study}
	%
	%
	% author names and IEEE memberships
	% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
	% a structure at a ~ so this keeps an author's name from being broken across
	% two lines.
	% use \thanks{} to gain access to the first footnote area
	% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
	% was not built to handle multiple paragraphs
	%

@@

	% As a general rule, do not put math, special symbols or citations
	% in the abstract or keywords.
	\begin{abstract}
		This paper proposes a deep convolutional neural network model for ordinal regression based on the use of a family of probabilistic ordinal link functions in the output layer called Cumulative Link Models. This kind of models projects each pattern into a 1-dimensional space. In this case, the projections are estimated by a non-linear deep neural network. A set of ordered thresholds splits this space into the different classes of the problem. Different link functions will be studied in this experimental study, and the results will be contrasted with statistical analysis. To further improve the results, we combine these ordinal models with a loss function that takes into account the distance between the categories, based on the weighted Kappa index. The experiments run over two different ordinal classification problems, and the statistical tests confirm that these models improve the results of a nominal model and outperform the proposals considered in the literature.
	\end{abstract}

	% Note that keywords are not normally used for peerreview papers.
	\begin{IEEEkeywords}
		Deep learning, Ordinal Regression, Link models.
	\end{IEEEkeywords}

@@

	Ordinal classification problems are those classification tasks where labels are ordered, and there are different inter-classes importances for each pair of classes. This kind of problem can be treated as a nominal classification problem, but doing this discards the ordinal information. A better approach is to use specific methods that take into account this kind of information to improve the performance of the classification model. The Proportional Odds Model (POM)~\cite{agresti2010analysis} is an ordinal alternative to the softmax function that was designed for ordinal regression. It belongs to a wider family of models called Cumulative Link Models (CLM). It is inspired in the concept of a latent variable that is projected in an n-dimensional space and a set of thresholds that divides that space into the different ordinal levels. This kind of models uses a link function which can be of different types. The most common link function is the \texttt{logit} one, that is used in POM. However, there are other functions that can be explored. This model will be described in depth in Section \ref{sect:ordinalproblem}.

	Ordinal models can also be applied to deep learning models. In the case of convolutional networks, the model projection used by the threshold model will be obtained from the last layer of the network. When working with a 1-dimensional space, the last layer will have only one neuron, and its value will be used to classify every sample in the corresponding class according to the thresholds. Some previous works have used the POM in traditional neural networks~\cite{gutierrez2016ordinal}, but it has not been applied to convolutional networks yet. Also, there are some link functions that has not been explored.

	In this paper, an experimental study regarding link functions for Cumulative Link Models in deep learning will be made. Also, other parameters that can affect the training process and the model performance, like the learning rate of the optimization algorithm, the batch size, and its interaction will be studied. The nominal version of this model will be used as a baseline for comparison. We will contrast the results obtained with statistical analysis to provide more robust conclusions. An approximated ANOVA III \cite{miller1997beyond} test followed by a posthoc Tukey's test \cite{tukey1949comparing} will be performed because of the limitations of the computational time required to run a higher number of executions.

	The experiments will be run using two different ordinal datasets: Diabetic Retinopathy~\cite{de2018weighted}, which contains high-resolution fundus images related with diabetes disease, and Adience~\cite{beckham2017unimodal}, which includes human faces images associated with an age range.

	This paper is organized as follows: in Section~\ref{sect:relatedwork}, we take a look at previous works related to this paper. Section~\ref{sect:ordinalproblem} present a formal description of an ordinal problem and the Cumulative Link Models. In Section~\ref{sect:experiments}, we describe the model, the experiments and the datasets used, while, in Section~\ref{sect:results}, we present the results obtained and the statistical analysis. Finally, Section~\ref{sect:conclusions} exposes the conclusions of this work.

	\section{Related works}
	\label{sect:relatedwork}
	There are many works related to the application and development of CNN models. However, few works are focused on ordinal classification problems.

	J. de la Torre et al.~\cite{de2018weighted} proposed the use of a continuous version of the QWK metric as loss function for the optimization algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces overfitting and training time. Also, they checked the importance of hyper-parameter tuning. First, they defined QWK metric as follows:
	\begin{equation}
	\text{QWK} = 1 - \frac{\sum\limits_{i,j} \omega_{i,j} O_{i,j}}{\sum\limits_{i,j} \omega_{i,j} E_{i,j}},
	\end{equation}
	where $\omega$ is the penalization matrix (in this case, quadratic weights are considered), $O$ is the confusion matrix and $E$ is the normalized outer product between the prediction and the true vector.

	Then, they provided a continuous version of the QWK metric ($\text{QWK}_c$) based on the probabilities of the predictions:
	\begin{equation}
	\text{QWK}_c = 1 - \frac{\sum\limits_{k=1}^N \sum\limits_{q=1}^Q \omega_{t_k, q} P(y = \mathcal{C}_q | \mathbf{x}_k)}{\sum\limits_{i=1}^Q \frac{N_i}{N} \sum\limits_{j=1}^Q ( \omega_{i,j} \sum\limits_{k=1}^N P(y = \mathcal{C}_j | \mathbf{x}_k))},
	\end{equation}
	where $\text{QWK}_c \in [-1,1]$, $\mathbf{x}_k$ and $t_k$ are the input data and the real class of the $k$-th sample, $Q$ is the number of classes, $N$ is the number of samples, $N_i$ is the number of samples of the $i$th class, $P(y = \mathcal{C}_q | \mathbf{x}_k)$ is the probability that the $k$th sample belongs to class $\mathcal{C}_q$ and $\omega_{i,j}$ are the elements of the penalization matrix. Generally, $\omega_{i,j} = \frac{|i-j|^n}{(C-1)^n}$, where $\omega_{i,j} \in [0,1]$.

	Then, they defined the loss function based on the QWK as follows:
	\begin{equation}
	\mathscr{L} = \log(1 - \text{QWK}_c),
	\end{equation}
	where $\mathscr{L} \in [-\infty, \log 2]$. This loss is a function to be minimized while the QWK metric must be maximized.

	Z. Niu et al. \cite{niu2016ordinal} proposed a learning approach to address ordinal regression problems using convolutional neural networks. They divided the problem into a series of binary classification sub-problems and proposed a multiple output CNN optimization algorithm to collectively resolve these classification sub-problems, taking into account the correlation between them.

	C. Beckham and C. Pal~\cite{beckham2017unimodal} proposed a straightforward technique to constrain discrete ordinal probability distributions to be unimodal, via the use of the Poisson and binomial probability distributions. They evaluated this approach in the context of deep learning on two large ordinal image datasets, including the Adience dataset used in this paper, obtaining promising results. Also, they proposed a simple squared-error reformulation \cite{beckham2016simple} that was sensitive to class ordering.

	Adience dataset has been used in other works for human age estimation. E. Eidinger \cite{eidinger2014age} presented an approach using support vector machines and neural networks. J.-C. Chen \cite{chen2016cascaded} proposed a coarse-to-fine strategy for deep convolutional networks. G. Levi \cite{levi2015age} presented another convolutional network model for age estimation. \textcolor{red}{Estos son ordinales?} \textcolor{blue}{No son ordinales pero los cito aqui porque luego los uso para comparar con ellos ya que son los que he encontrado que utilicen las base de datos Adience para predecir edad.}

	H. Li et al. \cite{li2017deep} applied deep learning techniques for solving the ordinal problem of Alzheimer's diagnosis and detecting the different levels of the disease.

	Y. Liu et al. \cite{liu2017deep} proposed a new approach of which transforms the ordinal regression problem to binary classification problems and use triplets with instances from different categories to train deep neural networks. In this way, high-level features describing the ordinal relationship are extracted automatically.

	A. Rios et al. \cite{rios2017ordinal} presented a CNN model designed to handle ordinal regression tasks on psychiatric notes. They combined an ordinal loss function, a CNN model and conventional feature engineering. Also, they applied a technique called Locally Interpretable Model-agnostic Explanation (LIME) to make the non-linear model more interpretable.

	S. Chen et al. \cite{chen2017using} proposed a deep method termed Ranking-CNN. This method combines multiple binary CNNs that are trained with ordinal age labels. The binary outputs are aggregated for the final age prediction and they achieved a tighter error bound for ranking-based age estimation.

	H. Fu et al. \cite{fu2018deep} applied deep learning techniques to Monocular Depth Estimation. They introduced a spacing-increasing discretization strategy to treat the problem as an ordinal regression problem. They improved the performance when training the network with an ordinary regression loss. Also, they used a multi-scale network structure that avoids unnecessary spatial pooling.

	Y. Liu et al. \cite{liu2018constrained} proposed a constrained optimization formulation for the ordinal regression problem which minimizes the negative loglikelihood for multiple categories constrained by the order relationship between instances.

	A. Pal et al. \cite{pal2018severity} defined a loss function for CNN that is based on the Earth Mover's Distance and takes into account the ordinal class relationships.

	M. ALALI et al. \cite{alali2018multi} proposed a complex CNN architecture for solving Twitter Sentiment Classification as an ordinal problem. They checked that using average pooling preserves significant features that provide more expresiveness to ordinal scale.

	\section{Cumulative Link Models (CLM)}
	\label{sect:ordinalproblem}
	An ordinal classification problem consists of predicting the label $y$ of an input vector $\mathbf{x}$, where $\mathbf{x} \in \mathcal{X} \subseteq \mathds{R}^K$ and $y \in \mathcal{Y}~=~\{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_Q\}$, i.e. $\mathbf{x}$ is in a K-dimensional input space, and $y$ is in a label space of $Q$ different labels. The objective of the ordinal problem is to find a function $r : \mathcal{X} \rightarrow \mathcal{Y}$ to predict the labels or categories of new patterns, given a training set of $N$ points, $D = \{(\mathbf{x}_i, y_i), i = 1, ..., N\}$. Labels have a natural ordering in ordinal problems: $\mathcal{C}_1 \prec \mathcal{C}_2 \prec ... \prec \mathcal{C}_Q$. The order between labels makes it possible to compare two different elements of $\mathcal{Y}$ by using the relation $\prec$. This is not possible under the nominal classification setting. In regression (where $y \in \mathds{R}$), real values in $\mathds{R}$ can be ordered by the standard $<$ operator, but labels in ordinal regression ($y \in \mathcal{Y}$) do not carry metric information, so the category serves as a qualitative indication of the pattern rather than a quantitative one.

	The Proportional Odds Model (POM) arises from a statistical background and is one of the first models designed explicitly for ordinal regression~\cite{mccullagh1980regression}. It dated back to 1980 and is a member of a wider family of models recognised as Cumulative Link Models (CLM)~\cite{agresti2010analysis}. CLMs predict probabilities of groups of contiguous categories, taking the ordinal scale into account. In this way, cumulative probabilities $P(y \prec C_q |\mathbf{x})$ are estimated, which can be directly related to standard probabilities:
	\begin{align}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) &= P(y = \mathcal{C}_1 | \mathbf{x}) + ... + P(y = \mathcal{C}_q | \mathbf{x}),\\
		P(y = \mathcal{C}_q | \mathbf{x}) &= P(y \preceq \mathcal{C}_q | \mathbf{x}) - P(y \preceq \mathcal{C}_{q-1} | \mathbf{x}),
	\end{align}
	with $q = 2, ..., Q-1$, and considering that $P(y = \mathcal{C}_1 | \mathbf{x}) = P(y \preceq \mathcal{C}_1 | \mathbf{x})$ and $P(y \preceq \mathcal{C}_Q | \mathbf{x}) = 1$.

	The model is inspired by the notion of a latent variable, where $f(\textbf{x})$ represents a one-dimensional mapping obtained from the output of the last layer, which has only one neuron. The decision rule $r: \mathcal{X} \rightarrow \mathcal{Y}$ is not fitted directly, but stochastic ordering of space $\mathcal{X}$ is satisfied by the following general model form \cite{herbrich2000large}:
	\begin{equation}
		g^{-1}(P(y \preceq \mathcal{C}_q | \mathbf{x})) = b_q - f(\mathbf{x}), \quad q = 1, ..., Q-1,
	\end{equation}
	where $g^{-1} : [0,1] \rightarrow (-\infty, +\infty)$ is a monotonic function often termed as the inverse link function, and $b_q$ is the threshold defined for class $\mathcal{C}_q$. Consider the latent variable $y^* = f(\mathbf{x})^* = f(\mathbf{x}) + \epsilon$, where $\epsilon$ is the random variable of the error. The most common choice for the probability distribution of $\epsilon$ is the logistic function (which is the default function for POM). Label $\mathcal{C}_q$ is predicted if and only if $f(\mathbf{x}) \in [b_{q-1}, b_q]$, where the function $f$ and $\mathbf{b} = (b_0, b_1, ..., b_{Q-1}, b_Q)$ are to be determined from the data. It is assumed that $b_0 = -\infty$ and $b_Q = +\infty$, so the real line defined by $f(\textbf{x}), \textbf{x} \in \mathcal{X}$, is divided into $Q$ consecutive intervals. Each interval corresponds to a category. The constraints $b_1 \le b_2 \le ... \le b_{Q-1}$ ensures that $P(y \preceq \mathcal{C}_q | \mathbf{x})$ increases with $q$ \cite{mccullagh1980regression}.
	This order is achieved by defining the first threshold and calculating the rest of thresholds from the first in the following form:

	\begin{equation}
	b_n = b_1 + \sum_{i=1}^{n-1} \alpha_i^2, \quad n = 2, ..., N,
	\end{equation}
	where $b_1$ and $\alpha_i$ are learnable parameters, and $N$ is the number of classes.

	In this work, we use this ordinal model with different link functions for the probability distribution of $\epsilon$, including \texttt{logit}, \texttt{probit} and complementary log-log (\texttt{clog-log}).

	These three types of links are explained below and represented in Figure \ref{fig:linkfunctions}. They all follow the same form $P(y \preceq \mathcal{C}_q | \mathbf{x}) = \Phi(b_q - f(\mathbf{x}))$ for a continuous cdf~$\Phi$.

	\begin{figure}[!t]
		\centering
		\includegraphics[width=3in]{img/linkfunctions.pdf}
		\caption{Representation of link functions.}
		\label{fig:linkfunctions}
	\end{figure}

	\begin{itemize}
		\item \texttt{Logit}. \texttt{logit} link function is the function used for the Proportional Odds Model. The \texttt{logit} link is:
		\begin{equation}
		\begin{aligned}
		\text{logit}[P(y \preceq \mathcal{C}_q | \mathbf{x})] = \log\frac{P(y \preceq \mathcal{C}_q | \mathbf{x})}{1 - P(y \preceq \mathcal{C}_q | \mathbf{x})}=& \\ = b_q - f(\mathbf{x}), \quad q = 1, ..., Q-1,
		\end{aligned}
		\label{eq:logit}
		\end{equation}		
		or the equivalent expression:		
		\begin{equation}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = \frac{1}{1 + e^{-(b - f(\mathbf{x}))}}.
		\label{eq:logit2}
		\end{equation}

		\item \texttt{Probit}. \texttt{probit} link function is the inverse of the standard normal cumulative distribution function (cdf) $\Phi$. Its expression is:
		\begin{equation}
		\begin{aligned}
		\Phi^{-1}[P(y \preceq \mathcal{C}_q | \mathbf{x})] = b_q - f(\mathbf{x}), \quad &q = 1, ..., Q-1,\\
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = \Phi(b_q - f(\mathbf{x})), \quad &q = 1, ..., Q-1,
		\end{aligned}
		\label{eq:probit}
		\end{equation}		
		which can also be expressed as:
		\begin{equation}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = \int_{-\infty}^{b_q - f(\mathbf{x})} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2}x^2} \mathrm{d}x.
		\label{eq:probit2}
		\end{equation}

		\item \textit{Complementary log-log}. Like the \texttt{logit} and the \texttt{probit} transformation, the complementary log-log transformation takes a response that is restricted to the $(0,1)$ interval and converts it into something in the $(-\infty, +\infty)$ interval. Complementary log-log expression is:
		\begin{equation}
		\begin{aligned}
		\log[-\log[1 - P(y \preceq \mathcal{C}_q | \mathbf{x})]] =&\\
		= b_q - f(\mathbf{x}), \quad q = 1, ..., Q-1,&
		\end{aligned}
		\label{eq:cloglog}
		\end{equation}
		that is:
		\begin{equation}
		P(y \preceq \mathcal{C}_q | \mathbf{x}) = 1 - e^{-e^{b_q - f(\mathbf{x})}}, \quad q = 1, ..., Q-1.
		\label{eq:cloglog2}
		\end{equation}
	\end{itemize}

>>>>

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***
% The testflow support page is at:                                        ***
% http://www.michaelshell.org/tex/testflow/                               ***

@@

\begin{document}
	%
	% paper title
	% Titles are generally capitalized except for words such as a, an, and, as,
	% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
	% not capitalized unless they are the first or last word of the title.
	% Linebreaks \\ can be used within to get better formatting as desired.
	% Do not put math or special symbols in the title.
	\title{Deep ordinal classification based on cumulative link models}
	%
	%
	% author names and IEEE memberships
	% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
	% a structure at a ~ so this keeps an author's name from being broken across
	% two lines.
	% use \thanks{} to gain access to the first footnote area
	% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
	% was not built to handle multiple paragraphs
	%

@@

	% As a general rule, do not put math, special symbols or citations
	% in the abstract or keywords.
	\begin{abstract}
		This paper proposes a deep convolutional neural network model for ordinal regression by considering a family of probabilistic ordinal link functions in the output layer. The link functions are those used for cumulative link models, which are traditional statistical linear models based on projecting each pattern into a 1-dimensional space. A set of ordered thresholds splits this space into the different classes of the problem. In our case, the projections are estimated by a non-linear deep neural network. To further improve the results, we combine these ordinal models with a loss function that takes into account the distance between the categories, based on the weighted Kappa index. Three different link functions are studied in the experimental study, and the results are contrasted with statistical analysis. The experiments run over two different ordinal classification problems and the statistical tests confirm that these models improve the results of a nominal model and outperform other proposals considered in the literature.
	\end{abstract}

	% Note that keywords are not normally used for peerreview papers.
	\begin{IEEEkeywords}
		Deep learning, ordinal regression, cumulative link models.
	\end{IEEEkeywords}

@@

	Ordinal classification problems are those classification tasks where labels are ordered, and there are different inter-classes importances for each pair of classes. This kind of problem can be treated as a nominal classification problem, but this discards the ordinal information. A better approach is to use specific methods that take the ordinality into account to improve the performance of the classification model. The Proportional Odds Model (POM)~\cite{mccullagh1980regression} is an ordinal alternative to the binary logistic regression. It belongs to a wider family of models called Cumulative Link Models (CLMs)~\cite{agresti2010analysis}. CLMs are inspired in the concept of a latent variable that is projected in a 1-dimensional space and a set of thresholds that divides that space into the different ordinal levels. This kind of models uses a link function which can be of different types, although the most common link function is the \texttt{logit}, which is used in POM. We explore different alternatives, as explained in depth in Section \ref{sect:ordinalproblem}.

	Ordinal models can also be applied to deep learning models. In the case of CNNs, the model projection used by the threshold model can be obtained from the last layer of the network. Given that we work with a 1-dimensional space, the last layer has only one neuron (projection of the pattern), and its value is used to classify the sample into the corresponding class according to the thresholds. Some previous works have used the \texttt{logit} in shallow neural networks~\cite{gutierrez2016ordinal}, but this strategy has not been considered for deep learning, and alternative link functions have not been evaluated. To further improve the results, we train these models by minimising an ordinal loss function based on the Weighted Kappa index \cite{de2018weighted}, instead of using the standard cross-entropy.

	In this paper, we propose the use of CLMs for deriving deep learning ordinal classifiers. An experimental study evaluating the three most common link functions is performed. Also, other parameters that can affect the training process and the model performance are studied, such as the learning rate of the optimization algorithm, the batch size, and their interaction. The nominal version of this model is used as a baseline for comparison. We contrast the results obtained with a statistical analysis to provide more robust conclusions. An ANOVA III test \cite{miller1997beyond} followed by a posthoc Tukey's test \cite{tukey1949comparing} is performed over $5$ runs of the experiments, because of the demands of computational time required to run a higher number of executions. The experiments are run using two different ordinal datasets: Diabetic Retinopathy~\cite{de2018weighted}, which contains high-resolution fundus images related with diabetes disease, and Adience~\cite{beckham2017unimodal}, which includes human faces images associated with an age range.

	This paper is organized as follows: in Section~\ref{sect:relatedwork}, we take a look at previous works related to this paper. Section~\ref{sect:ordinalproblem} present a formal description of an ordinal problem and CLMs. In Section~\ref{sect:experiments}, we describe the model, the experiments and the datasets used, while, in Section~\ref{sect:results}, we present the results obtained and the statistical analysis. Finally, Section~\ref{sect:conclusions} exposes the conclusions of this work.

	\section{Related works}
	\label{sect:relatedwork}
	There are many works related to the application and development of CNN models \cite{liu2017deep}, but few works focus on ordinal classification problems. The approaches proposed on those works are mainly related to 
	solving the ordinal problem as multiple binary sub-problems, using an ordinal loss function, solving the problem as a multi-class problem with constraints or simply using an ordinal evaluation metric. These works are described below:

	\vspace{0.2cm}
	$\bullet$ \textit{Solving the ordinal problem as multiple binary sub-problems.}

%		\textbf{(Binarización. Resolución de varios subproblemas.)}
	Z. Niu et al. \cite{niu2016ordinal} proposed a learning approach to address ordinal regression problems using convolutional neural networks. They divided the problem into a series of binary classification sub-problems and proposed a multiple output CNN optimization algorithm to collectively solve these classification sub-problems, taking into account the correlation between them.

%		\textbf{(Binarización. Resolución de varios subproblemas.)}
	H. Li et al. \cite{li2017deep} applied deep learning techniques for solving the ordinal problem of Alzheimer's diagnosis and detecting the different levels of the disease as multiple binary sub-problems.

%		\textbf{(Binarización. Resolución de varios subproblemas.)}
	Y. Liu et al. \cite{liu2017deep} proposed a new approach which transforms the ordinal regression problem to binary classification sub-problems and use triplets with instances from different categories to train deep neural networks. In this way, high-level features describing the ordinal relationship are extracted automatically. Given that triplets must be generated, this approach is only recommended for small datasets.

%		\textbf{(Binarización. Resolución de varios subproblemas.)}
	S. Chen et al. \cite{chen2017using} proposed a deep learning method termed Ranking-CNN. This method combines multiple binary CNNs that are trained with ordinal age labels. The binary outputs are aggregated for the final age prediction. They achieved a tighter error bound for ranking-based age estimation.

	\vspace{0.2cm}
	$\bullet$ \textit{Using an ordinal loss function.}

%		\textbf{(Función de coste ordinal)}
	J. de la Torre et al.~\cite{de2018weighted} proposed the use of a continuous version of the QWK metric as loss function for the optimization algorithm. They compared this cost function against the traditional log-loss function across three different databases, including the Diabetic Retinopathy database as the most complex one. They proved that their function could improve the results as it reduces overfitting and training time. Also, they checked the importance of hyper-parameter tuning.

%		\textbf{(Función de coste ordinal y distribución unimodal.)}
	C. Beckham and C. Pal~\cite{beckham2017unimodal} proposed a straightforward technique to constrain discrete ordinal probability distributions to be unimodal, via the use of the Poisson and binomial probability distributions. The parameters of these distributions were learnt by using a deep neural network. They evaluated this approach on two large ordinal image datasets, including the Adience dataset used in this paper, obtaining promising results. They also proposed a simple squared-error reformulation \cite{beckham2016simple} that was sensitive to class ordering.

%		\textbf{(Funcion de coste ordinal)}
	A. Rios et al. \cite{rios2017ordinal} presented a CNN model designed to handle ordinal regression tasks on psychiatric notes. They combined an ordinal loss function, a CNN model and conventional feature engineering. Also, they applied a technique called Locally Interpretable Model-agnostic Explanation (LIME) to make the non-linear model more interpretable.

%		\textbf{(Funcion de coste ordinal)}
	H. Fu et al. \cite{fu2018deep} applied deep learning techniques to Monocular Depth Estimation. They introduced a spacing-increasing discretization strategy to treat the problem as an ordinal regression problem. They improved the performance when training the network with an ordinal regression loss. Also, they used a multi-scale network structure that avoids unnecessary spatial pooling.
%	\textcolor{red}{La parte ordinal no queda muy clara}

%		\textbf{(Funcion de coste ordinal)}
	A. Pal et al. \cite{pal2018severity} defined a loss function for CNN that is based on the Earth Mover's Distance and takes into account the ordinal class relationships.

	\vspace{0.2cm}
	$\bullet$ \textit{Solving the problem as a multi-class problem with constraints.}

%		\textbf{(Multi-clase con restricciones)}
	Y. Liu et al. \cite{liu2018constrained} proposed a constrained optimization formulation for the ordinal regression problem which minimizes the negative loglikelihood for a multi-class problem constrained by the order relationship between instances.

	\vspace{0.2cm}
	$\bullet$ \textit{Simply using an ordinal evaluation metric.}

%		\textbf{(Métrica ordinal (MAE))}
	M. ALALI et al. \cite{alali2018multi} proposed a complex CNN architecture for solving Twitter Sentiment Classification as an ordinal problem. They checked that using average pooling preserves significant features that provide more expressiveness to ordinal scale. They didn't propose any method to include the ordinal information into the classifier but they tried to find the best CNN model architecture based on an ordinal metric.

	Adience dataset has been used in other works for human age estimation but most of them solved the problem as a multi-class problem instead of using the ordinal relation between classes. E. Eidinger \cite{eidinger2014age} presented an approach using support vector machines and neural networks. J.-C. Chen \cite{chen2016cascaded} proposed a coarse-to-fine strategy for deep CNNs, while G. Levi \cite{levi2015age} presented another convolutional network model for age estimation.

	However, none of the works mentioned have combined a CNN model with an unimodal probability distribution, without dividing the problem in multiple binary sub-problems and using a continuous loss function based on one of the most common metrics for ordinal classification (Weighted Kappa). This is the approach that will be proposed in this work for solving two separate ordinal problems.

%	The QWK metric is defined as follows:
%	\begin{equation}
%	\nonumber
%	\text{QWK} = 1 - \frac{\sum\limits^N_{i,j} \omega_{i,j} O_{i,j}}{\sum\limits^N_{i,j} \omega_{i,j} E_{i,j}},
%	\end{equation}
%	where $N$ is the number of samples rated, $\omega$ is the penalization matrix (in this case, quadratic weights are considered), $O$ is the confusion matrix, $E_{ij} = \frac{O_{i\bullet} O_{\bullet j}}{N}$, $O_{i\bullet}$ is the sum of the $i\text{-th}$ row and $O_{\bullet j}$ is the sum of the $j\text{-th}$ column.
%	
%	Then, they provided a continuous version of the QWK metric ($\text{QWK}_c$) based on the probabilities of the predictions:
%	\begin{equation}
%	\nonumber
%	\text{QWK}_c = 1 - \frac{\sum\limits_{k=1}^N \sum\limits_{q=1}^Q \omega_{t_k, q} P(y = \mathcal{C}_q | \mathbf{x}_k)}{\sum\limits_{i=1}^Q \frac{N_i}{N} \sum\limits_{j=1}^Q \left( \omega_{i,j} \sum\limits_{k=1}^N P(y = \mathcal{C}_j | \mathbf{x}_k)\right)},
%	\end{equation}
%	where $\text{QWK}_c \in [-1,1]$, $\mathbf{x}_k$ and $t_k$ are the input vector and the real class of the $k$-th sample, respectively, $Q$ is the number of classes, $N$ is the number of samples, $N_i$ is the number of samples of the $i$-th class, $P(y = \mathcal{C}_q | \mathbf{x}_k)$ is the probability that the $k$-th sample belongs to class $\mathcal{C}_q$ and $\omega_{i,j}$ are the elements of the penalization matrix. Generally, $\omega_{i,j} = \frac{|i-j|^n}{(C-1)^n}$, where $\omega_{i,j} \in [0,1]$.
%	
%	Then, they defined the loss function based on the QWK as follows:
%	\begin{equation}
%	\nonumber
%	\mathscr{L} = \log(1 - \text{QWK}_c),
%	\end{equation}
%	where $\mathscr{L} \in [-\infty, \log 2]$. $\mathscr{L}$ should be minimized, while the $\text{QWK}_c$ metric should be maximized.

%	\textcolor{red}{Yo intentaría organizar un poco el análisis que has hecho. Por ejemplo, podrías distinguir entre trabajos que modifican la función de error o los que modifican el modelo (ya sea aplicando binarización o utilizando distribuciones unimodales). o los que hacen las dos cosas. También habría que ir poniendo posibles pegas a cada artículo y destacar las ventajas de lo que proponemos (se impone una distribución unimodal, sin aplicar binarización, junco con una versión continua de una de las funciones d eerror más usadas)}

	\section{Cumulative Link Models (CLM)}
	\label{sect:ordinalproblem}
	An ordinal classification problem consists in predicting the label $y$ of an input vector $\mathbf{x}$, where $\mathbf{x} \in \mathcal{X} \subseteq \mathds{R}^K$ and $y \in \mathcal{Y}~=~\{\mathcal{C}_1, \mathcal{C}_2, ..., \mathcal{C}_Q\}$, i.e. $\mathbf{x}$ is in a K-dimensional input space, and $y$ is in a label space of $Q$ different labels. The objective of the ordinal problem is to find a function $r : \mathcal{X} \rightarrow \mathcal{Y}$ to predict the labels or categories of new patterns, given a training set of $N$ samples, $D = \{(\mathbf{x}_i, y_i), i = 1, ..., N\}$. Labels have a natural ordering in ordinal problems: $\mathcal{C}_1 \prec \mathcal{C}_2 \prec ... \prec \mathcal{C}_Q$. The order between labels gives us the possibility to compare two different elements of $\mathcal{Y}$ by using the relation $\prec$. This is not possible under the nominal classification setting. In regression (where $y \in \mathds{R}$), real values in $\mathds{R}$ can be ordered by the standard $<$ operator, but labels in ordinal regression ($y \in \mathcal{Y}$) do not carry metric information, i.e. the category serves as a qualitative indication of the pattern rather than a quantitative one.

	The Proportional Odds Model (POM) arises from a statistical background and is one of the first models designed explicitly for ordinal regression~\cite{mccullagh1980regression}. It dated back to 1980 and is a member of a wider family of models recognised as Cumulative Link Models (CLMs)~\cite{agresti2010analysis}. CLMs predict probabilities of groups of contiguous categories, taking the ordinal scale into account. In this way, cumulative probabilities $P(y \prec C_q |\mathbf{x})$ are estimated, which can be directly related to standard probabilities:
	\begin{align}
		\nonumber P(y \preceq \mathcal{C}_q | \mathbf{x}) &= P(y = \mathcal{C}_1 | \mathbf{x}) + ... + P(y = \mathcal{C}_q | \mathbf{x}),\\
		\nonumber P(y = \mathcal{C}_q | \mathbf{x}) &= P(y \preceq \mathcal{C}_q | \mathbf{x}) - P(y \preceq \mathcal{C}_{q-1} | \mathbf{x}),
	\end{align}
	with $q = 2, ..., Q-1$, and considering that $P(y = \mathcal{C}_1 | \mathbf{x}) = P(y \preceq \mathcal{C}_1 | \mathbf{x})$ and $P(y \preceq \mathcal{C}_Q | \mathbf{x}) = 1$.

	The model is inspired by the notion of a latent variable, where $f(\textbf{x})$ represents a one-dimensional mapping. The decision rule $r: \mathcal{X} \rightarrow \mathcal{Y}$ is not fitted directly, but stochastic ordering of space $\mathcal{X}$ is satisfied by the following general model form \cite{herbrich2000large}:
	\begin{equation}
		\nonumber
		g^{-1}(P(y \preceq \mathcal{C}_q | \mathbf{x})) = b_q - f(\mathbf{x}), \quad q = 1, ..., Q-1,
	\end{equation}
	where $g^{-1} : [0,1] \rightarrow (-\infty, +\infty)$ is a monotonic function often termed as the inverse link function, and $b_q$ is the threshold defined for class $\mathcal{C}_q$. Consider the latent variable $y^* = f(\mathbf{x})^* = f(\mathbf{x}) + \epsilon$, where $\epsilon$ is the random component of the error. The most common choice for the probability distribution of $\epsilon$ is the logistic function (which is the default function for POM). Label $\mathcal{C}_q$ is predicted if and only if $f(\mathbf{x}) \in [b_{q-1}, b_q]$, where the function $f$ and $\mathbf{b} = (b_0, b_1, ..., b_{Q-1}, b_Q)$ are to be determined from the data. It is assumed that $b_0 = -\infty$ and $b_Q = +\infty$, so the real line defined by $f(\textbf{x}), \textbf{x} \in \mathcal{X}$, is divided into $Q$ consecutive intervals. Each interval corresponds to a category. The constraints $b_1 \le b_2 \le ... \le b_{Q-1}$ ensures that $P(y \preceq \mathcal{C}_q | \mathbf{x})$ increases with $q$ \cite{mccullagh1980regression}.

	In this work, we consider different link functions previously proposed in CLMs for the probability distribution of $\epsilon$, including \texttt{logit}, \texttt{probit} and complementary log-log (\texttt{clog-log}). These three types of links are explained below and represented in Figure \ref{fig:linkfunctions}. They all follow the same form $\text{link}[P(y \preceq \mathcal{C}_q | \mathbf{x})] = b_q - f(\mathbf{x})$.

	\begin{figure}[!t]
		\centering
		\includegraphics[width=3in]{img/linkfunctions.pdf}
		\caption{Different link functions commonly used for CLMs.}
		\label{fig:linkfunctions}
	\end{figure}
